---
title: "Midterm Project"
author: "Megan Tran"
date: 'October 23, 2022'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
library(ggplot2)
library(dtplyr)
library(dplyr)
library(stringr)
```


#Introduction 

The Contraceptive Care - All Women measure (CCW), as part of the Maternal and Infant Health Initiative, Contraceptive Care Quality grant, was compiled data taken from women ages 15-44 at risk for unintended pregnancy. The women were stratified in two age groups, those who are 15-20 and those who are 21-44. Contraceptive type either fell under the category of long-acting reversible methods of contraception (LARC), such as birth control implants or IUDs, or most/moderately effective methods of contraceptions (M/M), such as oral pills, patches, rings, injectables, female sterilization, or diaphragms. Data was gathered in the U.S. for 3 consecutive years, 2014-2016. Furthermore, "Rate of contraceptive use" in the dataset refers to those who use a type of contraceptive divided by those who are "eligible", or at risk for pregnancy.
 
LARC contraceptives are more effective forms of birth control than M/M methods and are safe for women of all ages to use. LARC methods could prevent more cases of teenage pregnancy for young women at risk and who are sexually active. However, M/M methods are considered to be more traditional and may be perceived by the public to be safer since they don't require surgical insertions of long-term devices into the body. Young women who have not received enough information about the different available contraceptive methods may not be able to make an informed decision on the contraception that's right for them.

The research question being explored is if younger women (age 15-20) use long-acting reversible methods of contraception (LARC) at a lower rate than older women (age 21-44) and is that trend consistent throughout the three year period?


#Methods 

The data was collected through administrative survey measures. The representative sample excluded U.S. women not at risk of unintended pregnancy because they were infecund for non-contraceptive reasons, had a live birth in the last 2 months of the measurement year, or were pregnant or their pregnancy outcome was unknown at the end of the year(s). Once the exclusions were applied, the sample included women who were not pregnant at any point in the 3-year period, those who had a live birth in the first 10 months of the measurement year(s), and those who had a miscarriage, stillbirth, ectopic pregnancy, or induced abortion. 


Reading in the data:
```{r read-data, cache=TRUE}
if (!file.exists("ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv")) {
download.file("https://data.chhs.ca.gov/dataset/c2698502-d276-4e55-9057-8153e39d21b1/resource/12a73f54-dcf4-4e38-843c-e988385be69b/download/ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv", "ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv", method="libcurl", timeout = 60) 
}
contra <- data.table::fread("ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv") 
```


First, the number of missing values were checked. 
```{r}
mean(is.na(contra))
```
There were no missing values in the dataset so there was no need to remove observations or impute data.


A regular expression was used in order to remove the "%" symbol from the "Rate of Contraceptive Use" column.
```{r}
contra$`Rate of Contraceptive Use` <- stringr::str_remove_all(contra$`Rate of Contraceptive Use`, "%")
```


Then the column was changed from a character variable to numeric.
```{r}
contra$`Rate of Contraceptive Use` <- as.numeric(contra$`Rate of Contraceptive Use`)
```


There's a wide range between the maximum and minimum rates.
```{r}
summary(contra$`Rate of Contraceptive Use`)
```


The average of the rates of contraceptive use was calculated by contraceptive type, age group, and year. The variable "avg_rate" was created.
```{r}
avg_contra <- contra[ , .(
    avg_rate = mean (`Rate of Contraceptive Use`)
  ), 
  by = .(`Contraceptive Type`, `Age Group`, Year)]

avg_contra
```

#Preliminary Results 

```{r}
avg_contra[!is.na(`avg_rate`)] %>% 
  ggplot()+
  geom_col(mapping=aes(x=`Contraceptive Type`, y= `avg_rate`, fill = `Contraceptive Type`)) +
   labs(title = "Bar Graphs of Contraceptive Use by Age Group") +
  labs( y = "Rate of Contraceptive Use") +
    facet_wrap( ~`Age Group`)
```
These bar graphs show the rates of contraceptive use, LARC or Most/Moderately Effective methods, per age group. The figure shows that older women use LARC methods at a slightly higher rate. Since these bar graphs are so similar, I want to see if the graphs look any different when I split them into years. Perhaps, a different trend will appear.


```{r}
avg_contra[!is.na(`avg_rate`)] %>% 
  ggplot()+
  geom_col(mapping=aes(x=`Contraceptive Type`, y= `avg_rate`, fill = `Contraceptive Type`)) +
   labs(title = "Bar Graphs of Contraceptive Use by Year and Age Group") +
   facet_wrap(Year ~ `Age Group`, nrow=3)
```
These bar graphs split the data by year. Interestingly enough, older women used LARC methods at a lower rate than younger women in 2014. In the next two years, you can observe that the rates for 15-20 year olds decreased. These graphs give us a better idea of how the rate distribution changes for the two age groups over the years. They suggest that time might be confounder in the association between contraceptive type and age group since the trend is different for different years. 



```{r}
contra[!is.na(`Rate of Contraceptive Use`)] %>% 
  ggplot()+
  geom_boxplot(mapping=aes(x=`Contraceptive Type`, y=`Rate of Contraceptive Use`, fill = `Contraceptive Type`)) +
   labs(title = "Boxplots of Contraceptive Use by Year and Age Group") + 
    facet_wrap(Year ~ `Age Group`, nrow=3)

```
Again, the boxplots depict the distribution of contraceptive rates by year, age group, and contraceptive type. Similar to the bar graphs, these boxplots confirm that the median LARC contraceptive use rates for younger women are only higher than for older women only in 2014. The opposite is true for the next two years. There seems to be an outlier for LARC contraceptive types in 2015 for 21-44 year olds. By examining the data more closely, we can see that the outlier is not particularly concerning because it comes from the "Other Asian/Pacific Islander" racial group and they consistently have higher rates throughout the dataset.




```{r}
database <- data.frame(
    Year = avg_contra$Year, 
   Contraceptive_Type = avg_contra$`Contraceptive Type`,
    Age_Group = avg_contra$`Age Group`,
   Average_Rate_of_Contraceptive_Use = avg_contra$avg_rate
 )
 knitr::kable(database, caption = "Contraceptive Use of Women 2014-2016")
```



#Conclusion 

Although statistical analyses should be conducted in order to determine a more accurate conclusion, we can look at the figures and table to reach a preliminary conclusion. In reference to the formulated research question, younger women do indeed use LARC contraceptive methods at a lower rate than older women, but this was not true for all 3 years, 2014-2016. In 2014, younger women actually had higher rates for LARC contraceptive use. More research should be conducted as to why there was a change between 2014 and 2015 and if the trend has held the same past 2016. Introducing LARC contraceptive methods to younger women could reduce unwanted pregnancies and offer alternative, long-term contraception options.













