---
title: "Midterm Project"
author: "Megan Tran"
date: 'October 23, 2022'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
library(ggplot2)
library(dtplyr)
library(dplyr)
```


```{r read-data, cache=TRUE}
if (!file.exists("ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv")) {
download.file("https://data.chhs.ca.gov/dataset/c2698502-d276-4e55-9057-8153e39d21b1/resource/12a73f54-dcf4-4e38-843c-e988385be69b/download/ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv", "ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv", method="libcurl", timeout = 60) 
}
contra <- data.table::fread("ofp-ccw-by-race-ethn_contra-type_age-group_14-16.csv") 
```


##Introduction 
(provide background on your dataset and formulated question)


The dataset was compiled for the Measure CCW, Contraceptive Care as part of the Maternal and Infant Health Initiative, Contraceptive Care Quality grant. 


Question: Are older women (age 21-44) more likely to use Long-Acting Reversible Contraceptives than younger women (age 15-20) and is the trend consistent across all races?




##Methods 
(include how and where the data were acquired, how you cleaned and wrangled the data, what tools you used for data exploration)



```{r}
mean(is.na(contra))
```

```{r}
str(contra)
```




##Preliminary Results 
(provide summary statistics in tabular form and publication-quality figures, take a look at the kable function from knitr to write nice tables in Rmarkdown)




##Conclusion about what you found in terms of the formulated question.














